\textbf{Preprinting Microbiology}

### 2017-02-08 submission/Schloss_PrePrints_mBio_2017.tex
Edit and strengthen Acknowledgements section

\textbf{\emph{Landscape of preprint servers.}} In 2013, two preprint
servers, the \emph{bioRxiv} (pronounced ``bio-archive'') and \emph{PeerJ
Preprints}, were launched as preprint servers for biologists that would
parallel \emph{arXiv} (8). Both platforms offer similar features:
preprint posting is free; each preprint receives a digital object
identifier (DOI) that facilitates the ability to cite preprints in other
scholarly work; if the preprint is ever published, the preprint is
linked to the published version; the submission process for both options
is relatively simple allowing authors to upload a PDF version of their
preprint and supplemental materials; preprints are typically publicly
available in under 24 hours; they have built in venues for authors to
discuss their research with people who leave comments on the manuscript;
preprints undergo a basic screening process to remove submissions with
offensive or non-scientific content; and the sites provide article-level
metrics indicating the number of times an abstract has been accessed or
the preprint has been downloaded. There are several important
differences between the two options. First, \emph{PeerJ Prints} is a
for-profit organization and \emph{bioRxiv} is a non-profit organization
sponsored by Cold Spring Harbor Laboratory. This difference can be
meaningful to authors since some journals, including the American
Society for Microbiology (ASM) Journals, will only accept submissions
that have been posted on preprint servers hosted by non-profit
organizations. Second, preprints at \emph{PeerJ Preprints} are posted
under the Creative Commons Attribution License (CC-BY) and
\emph{bioRxiv} preprints can be posted under one of four CC-BY licenses
or with no permission for reuse. This can be relevant for authors hoping
to submit their work to a journal as journals will not consider
manuscripts posted as preprints under a CC-BY license (e.g.
\emph{Proceedings of the National Academy of Sciences}). A cosmetic, but
still relevant difference between the two is the layout and feel of the
two websites. Compared to the \emph{bioRxiv} site, the \emph{PeerJ
Preprint} site is more fluid, gives readers the ability to ``follow'' a
preprint, and provides better access to article keywords and the ability
to search preprints. With broader acceptance of preprints by traditional
journals, many journals, including all of the ASM journals, have
established mechanisms to directly submit manuscripts that are posted as
preprints on \emph{bioRxiv}. It is only possible to transfer a
\emph{PeerJ Preprint} for submission to \emph{PeerJ}. In many ways,
preprint servers have taken on the feel of a journal. As adoption of
this approach to disseminating research expands, it is likely that the
features of these sites will continue to improve. It is also likely that
interfaces from third-parties will improve. For example, although Google
Scholar includes preprints hosted at \emph{bioRxiv} and \emph{PeerJ
Preprints} in their search results, PubMed and Web of Science do not.
There is also hope that the National Institutes of Health (NIH) will
renew their interest in including preprints in PubMed search results. As
preprint servers begin to look and act like traditional journals by
incorporating features and interfaces, it will be important to balance
the requirements placed on authors for those features with the current
efficiency of the preprint format.

@@

\textbf{\emph{Preprints have the potential to change the advancement of
science.}} Preprints are often viewed as existing in a state of
scientific limbo. As noted above, they represent a formal communication,
but are not officially published. As the use of preprints grows and
scientists' perceptions of preprints matures, there are a number of
issues that will need to be addressed. First, a common concern is that
if a researcher posts their work as a preprint, it will be ``scooped''
by another researcher and the preprint author will lose their ability to
claim primacy or their ability to publish the work in a journal.
Considering the preprint is a citable work with a DOI, it would, in
fact, be the preprint author that scooped the second. A growing number
of scientific societies and journals, including ASM view preprints as
citable and as having a legitimate claim to primacy (1, 14--16). Some
scientists worry that with such protection a researcher can make a claim
without valid data to support their claims (3). This is possible;
however, it is also the responsibility of the scientific community to
utilize the peer-review mechanisms that are available to comment on
those preprints pointing out methodological problems or to indicate that
they are speaking beyond the data. A second area of concern is whether a
preprint can be used to support a grant proposal. Given the length
limitations placed on grant proposals by funding agencies, there is a
push to cite previous work to indicate a research team's competence in
an area or to provide preliminary data. Some fear that the use of
preprints will allow scientists to circumvent page limits by posting
preliminary manuscripts. We would hope that both consumers of preprints
and grant proposal reviewers would be able to differentiate between
someone trying to game the system and someone that is using preprints as
a mechanism to improve their science. A third concern is what role
preprints should have in assessing a scientist's productivity. Clearly
use of publication metrics as an indicator of a scientist's productivity
and impact is a contentious topic without even discussing the role of
preprints. Regardless, given the propensity for researchers to list
manuscripts as being ``in preparation'' or ``in review'' on an
application or curriculum vitae, listing them instead as preprints that
can be reviewed by a committee would significantly enhance an
application and a reviewer's ability to judge the application. In fact,
several funding agencies including the Wellcome Trust and the UK Medical
Research Council encouraging fellowship applicants to include preprints
in their materials; meanwhile, the NIH is in the process of soliciting
input from the scientific community on their role in grant applications.
Others are mandating that researchers post preprints for all of their
work prior to submitting the work to a journal (17). Beyond these
concerns, preprints are also causing some to change their publication
goals. Some authors are explicitly stating that a preprint will not be
submitted to a journal (18). Although these authors may be a minority of
those who post preprints, such an approach may be attractive to those
that want to have a mechanism for people to cite a report of a brief
research communication, a critique of another publication, or negative
results. It is clear that the adoption of preprints will challenge how
scientists interact and evaluate each other's work. There is great
potential to empower researchers by controlling when a citable piece of
work is made public.

\textbf{\emph{Microbiology anecdotes.}} The peer-review editorial
process can be lengthy and adversarial. In contrast, preprints represent
a rapid and potentially collaborative method for disseminating research.
Several anecdotes from the microbiology literature are emblematic of
benefits of the rapid release cycle that is inherent in the use of
preprints. First, preprints have proven useful for rapidly disseminating
results for disease outbreaks and new technologies. Prior to the recent
Zika virus outbreak there were approximately 50 publications that
touched on the biology and epidemiology of the virus; as of January 2017
the number of Zika virus-related publications was over 1,700. During the
recent outbreak, more than 110 Zika virus-related preprints have been
posted at \emph{bioRxiv}. Any manuscript that was published went through
several month delays in releasing information to health care workers,
the public, and scientists needing to learn new methods to study a
previously obscure virus. In contrast, those that posted their work as a
preprint were able to disseminate their methods and results instantly.
Over the last several years there have also been rapid advances in DNA
sequencing technologies have fundamentally changed how microbial science
is performed. One notable technology, the MinIon sequencing platform
from Oxford Nanopore, has received considerable attention from
researchers who have posted more than 90 preprints describing new
MinIon-based methods and results to preprint servers. For such a rapidly
developing technology, the ability to share and consume methods from
other scientists has created a feed forward effect where the technology
has likely advanced at a faster rate than it otherwise would have.
Preprints have proven useful for rapidly correcting the scientific
literature. On February 9, 2015, Cell Systems posted a study by
Afshinnekoo et al. online (19). The study collected and analyzed
metagenomic sequence data from the New York City subway system and
reported finding \emph{Yersinia pestis} and \emph{Bacillus anthracis}.
Because of the focus on these two bioterrorism agents, this study
generated a considerable amount of media attention. On April 25, 2015,
Petit et al. (20) posted a preprint to Zenodo demonstrating that there
was no evidence for \emph{B. anthracis} in the dataset. On July 29,
2015, a critique was published by \emph{Cell Systems} along with a
response from the original authors offering a correction to their
manuscript (21, 22). A second anecdote of using preprints to aid in
post-publication peer-review surrounds the publishing of a draft
tardigrade genome in \emph{The Proceedings of the National Academy of
Sciences}. On November 23, 2015 a study by Boothby et al. (23) was first
published online. The authors posited that 17.5\% of its genes came from
bacteria, archaea, fungi, plants, and viruses. Another group had been
analyzing sequence data from a parallel tardigrade genome sequencing
project and did not observe the same result. A week later, on December
1, 2015, the second group had posted a preprint comparing the two genome
sequences and demonstrating that the exciting claims of horizontal gene
transfer were really the product of contaminants (24); this analysis
would eventually be published online by the original journal on March
24, 2016 followed by a rebuttal by the original authors on May 31, 2016
(25, 26). Two other analyses of the original data were published in May
2016 and a third was posted as a preprint on February 2, 2016 (27--29).
These anecdotes underscore the value of having a rapid posting cycle to
correcting errors in the scientific literature and that results posted
to preprint servers were able to correct the record within weeks of the
initial publication while the traditional path took six months in both
cases. A final notable case where preprints have accelerated the
correction of the scientific record was a preprint posted by Bik et al.
reporting numerous cases of image manipulation in peer reviewed studies
(30). Their preprint was posted on April 20, 2016 and published in
\emph{mBio} on June 7, 2016 (31). Instead of using preprints to react to
published papers that have been through peer review, it would be
interesting to consider how the editorial process for these examples and
the infamous ``Arsenic Life'' paper (32) would have been different had
they initially been posted as preprints.

@@

As the total number of preprints has grown exponentially since the
creation of \emph{bioRxiv}, submission of microbiology-affiliated
preprints has largely followed this growth (\textbf{Figure 1A}).
Although preprints are still relatively new, the collection of
microbiology-affiliated preprints indicates widespread experimentation
with the format and considerable geographic diversity. Reflecting the
still relatively novelty of preprints, 1,132 (86.1\%) corresponding
authors who submitted a microbiology-affiliated manuscript (N=1,314
total) have posted a single preprint and 3.6\% have posted 3 or more
preprints. Corresponding authors that have posted
microbiology-affiliated preprints are from 60 countries and are
primarily affiliated with institutions in the United States (50.8\% of
microbiology-affiliated preprints), United Kingdom (11.9\%), and Germany
(4.2\%). As the preprint format matures, it will be interesting to see
whether the fraction of authors that post multiple preprints increases
and whether the geographic diversity amongst those authors is
maintained.

As stated above, preprints offer researchers the opportunity to improve
the quality of their work by adding a more formal and public step to the
scientific process. Among the microbiology-affiliated manuscripts, 146
(9.3\%) had been commented on at least once and only 35 (2.2\%) more
than three times using the \emph{bioRxiv}-hosted commenting feature.
Although the hosted commenting is only one mechanism for peer review,
this result was somewhat disturbing since the preprint model implicitly
depends on people's willingness to offer others feedback. Regardless,
authors do appear to be incorporating feedback from colleagues or
editorial insights from journals as 404 (25.8\%) of preprints were
revised at least once. Among the preprints posted prior to January 1,
2016, 31.6\% of the Microbiology category preprints, 35.1\% of the
microbiology-affiliated preprints, and 33.8\% of all preprints have been
published. As noted above, not all authors submit their preprints to
journals. This would indicate that the ``acceptance rates'' are actually
higher. Regardless, considering that these acceptance rates are higher
than many peer-reviewed journals (e.g.~approximately 20\% at ASM
Journals), these results dispel the critique that preprints represent
overly preliminary research.

@@

This work was supported in part by funding to PDS from the National
Institutes of Health (P30DK034933). We appreciate the support of
Altmetric, Inc and Thompson Reuters who provided advanced programming
interface (API) access to their databases. The data, code, and an
executable R-markdown version of this manuscript are available at
\url{http://www.github.com/SchlossLab/Schloss_PrePrints_mBio_2017}.

>>>>

\textbf{\emph{Landscape of preprint servers.}} In 2013, two preprint
servers, the \emph{bioRxiv} (pronounced ``bio-archive'') and \emph{PeerJ
Preprints}, were launched as preprint servers for biologists that would
parallel \emph{arXiv} (8). Both platforms offer similar features:
preprint posting is free; each preprint receives a digital object
identifier (DOI) that facilitates the ability to cite preprints in other
scholarly work; if the preprint is ever published, the preprint is
linked to the published version; the submission process for both options
is relatively simple allowing authors to upload a PDF version of their
preprint and supplemental materials; preprints are typically publicly
available in under 24 hours; they have built in venues for authors to
discuss their research with people who leave comments on the manuscript;
preprints undergo a basic screening process to remove submissions with
offensive or non-scientific content; and the sites provide article-level
metrics indicating the number of times an abstract has been accessed or
the preprint has been downloaded. There are several important
differences between the two options. First, \emph{PeerJ Prints} is a
for-profit organization and \emph{bioRxiv} is a non-profit organization
sponsored by Cold Spring Harbor Laboratory. This difference can be
meaningful to authors since some journals, including the American
Society for Microbiology (ASM) Journals, will only accept submissions
that have been posted on preprint servers hosted by non-profit
organizations. Second, preprints at \emph{PeerJ Preprints} are posted
under the Creative Commons Attribution License (CC-BY) and
\emph{bioRxiv} preprints can be posted under one of four CC-BY licenses
or with no permission for reuse. This can be relevant for authors hoping
to submit their work to a journal as journals will not consider
manuscripts posted as preprints under a CC-BY license (e.g.
\emph{Proceedings of the National Academy of Sciences}). A cosmetic, but
still relevant difference between the two is the layout and feel of the
two websites. Compared to the \emph{bioRxiv} site, the \emph{PeerJ
Preprint} site is more fluid, gives readers the ability to ``follow'' a
preprint, and provides better access to article keywords and the ability
to search preprints. With broader acceptance of preprints by traditional
journals, many journals, including all of the ASM journals, have
established mechanisms to directly submit manuscripts that are posted as
preprints on \emph{bioRxiv}. It is only possible to transfer a
\emph{PeerJ Preprint} for submission to \emph{PeerJ}. In many ways,
preprint servers have taken on the feel of a journal. As adoption of
this approach expands, it is likely that the features of these sites
will continue to improve. It is also likely that interfaces from
third-parties will improve. For example, although Google Scholar
includes preprints hosted at \emph{bioRxiv} and \emph{PeerJ Preprints}
in their search results, PubMed and Web of Science do not. There is also
hope that the National Institutes of Health (NIH) will renew their
interest in including preprints in PubMed search results. As preprint
servers begin to look and act like traditional journals by incorporating
features and interfaces, it will be important to balance the
requirements placed on authors for those features with the current
efficiency of the preprint format.

@@

\textbf{\emph{Preprints have the potential to change the advancement of
science.}} Preprints are often viewed as existing in a state of
scientific limbo. As noted above, they represent a formal communication,
but are not officially published. As the use of preprints grows and
scientists' perceptions of preprints matures, there are a number of
issues that will need to be addressed.

First, a common concern is that if a researcher posts their work as a
preprint, it will be ``scooped'' by another researcher and the preprint
author will lose their ability to claim primacy or their ability to
publish the work in a journal. Considering the preprint is a citable
work with a DOI, it would, in fact, be the preprint author that scooped
the second. A growing number of scientific societies and journals,
including ASM view preprints as citable and as having a legitimate claim
to primacy (1, 14--16). Some scientists worry that with such protection
a researcher can make a claim without valid data to support their claims
(3). This is possible; however, it is also the responsibility of the
scientific community to utilize the peer-review mechanisms that are
available to comment on those preprints pointing out methodological
problems or to indicate that they are speaking beyond the data.

A second area of concern is whether a preprint can be used to support a
grant proposal. Given the length limitations placed on grant proposals
by funding agencies, there is a push to cite previous work to indicate a
research team's competence in an area or to provide preliminary data.
Some fear that the use of preprints will allow scientists to circumvent
page limits by posting preliminary manuscripts. We would hope that both
consumers of preprints and grant proposal reviewers would be able to
differentiate between someone trying to game the system and someone that
is using preprints as a mechanism to improve their science.

A third concern is what role preprints should have in assessing a
scientist's productivity. Clearly use of publication metrics as an
indicator of a scientist's productivity and impact is a contentious
topic without even discussing the role of preprints. Regardless, given
the propensity for researchers to list manuscripts as being ``in
preparation'' or ``in review'' on an application or curriculum vitae,
listing them instead as preprints that can be reviewed by a committee
would significantly enhance an application and a reviewer's ability to
judge the application. In fact, several funding agencies including the
Wellcome Trust and the UK Medical Research Council encouraging
fellowship applicants to include preprints in their materials;
meanwhile, the NIH is in the process of soliciting input from the
scientific community on their role in grant applications. Others are
mandating that researchers post preprints for all of their work prior to
submitting the work to a journal (17).

Beyond these concerns, preprints are also causing some to change their
publication goals. Some authors are explicitly stating that a preprint
will not be submitted to a journal (18). Although these authors may be a
minority of those who post preprints, such an approach may be attractive
to those that want to have a mechanism for people to cite a report of a
brief research communication, a critique of another publication, or
negative results. It is clear that the adoption of preprints will
challenge how scientists interact and evaluate each other's work. There
is great potential to empower researchers by controlling when a citable
piece of work is made public.

\textbf{\emph{Microbiology anecdotes.}} The peer-review editorial
process can be lengthy and adversarial. In contrast, preprints represent
a rapid and potentially collaborative method for disseminating research.
Several anecdotes from the microbiology literature are emblematic of
benefits of the rapid release cycle that is inherent in the use of
preprints.

First, preprints have proven useful for rapidly disseminating results
for disease outbreaks and new technologies. Prior to the recent Zika
virus outbreak there were approximately 50 publications that touched on
the biology and epidemiology of the virus; as of January 2017 the number
of Zika virus-related publications was over 1,700. During the recent
outbreak, more than 110 Zika virus-related preprints have been posted at
\emph{bioRxiv}. Any manuscript that was published went through several
month delays in releasing information to health care workers, the
public, and scientists needing to learn new methods to study a
previously obscure virus. In contrast, those that posted their work as a
preprint were able to disseminate their methods and results instantly.
Over the last several years there have also been rapid advances in DNA
sequencing technologies have fundamentally changed how microbial science
is performed. One notable technology, the MinIon sequencing platform
from Oxford Nanopore, has received considerable attention from
researchers who have posted more than 90 preprints describing new
MinIon-based methods and results to preprint servers. For such a rapidly
developing technology, the ability to share and consume methods from
other scientists has created a feed forward effect where the technology
has likely advanced at a faster rate than it otherwise would have.

Second, preprints have proven useful for rapidly correcting the
scientific literature. On February 9, 2015, \emph{Cell Systems} posted a
study by Afshinnekoo et al. online (19). The study collected and
analyzed metagenomic sequence data from the New York City subway system
and reported finding \emph{Yersinia pestis} and \emph{Bacillus
anthracis}. Because of the focus on these two bioterrorism agents, this
study generated a considerable amount of media attention. On April 25,
2015, Petit et al. (20) posted a preprint to Zenodo demonstrating that
there was no evidence for \emph{B. anthracis} in the dataset. On July
29, 2015, a critique was published by \emph{Cell Systems} along with a
response from the original authors offering a correction to their
manuscript (21, 22). A second anecdote of using preprints to aid in
post-publication peer-review surrounds the publishing of a draft
tardigrade genome in \emph{The Proceedings of the National Academy of
Sciences}. On November 23, 2015 a study by Boothby et al. (23) was first
published online. The authors posited that 17.5\% of its genes came from
bacteria, archaea, fungi, plants, and viruses. Another group had been
analyzing sequence data from a parallel tardigrade genome sequencing
project and did not observe the same result. A week later, on December
1, 2015, the second group had posted a preprint comparing the two genome
sequences and demonstrating that the exciting claims of horizontal gene
transfer were really the product of contaminants (24); this analysis
would eventually be published online by the original journal on March
24, 2016 followed by a rebuttal by the original authors on May 31, 2016
(25, 26). Two other analyses of the original data were published in May
2016 and a third was posted as a preprint on February 2, 2016 (27--29).
These anecdotes underscore the value of having a rapid posting cycle to
correcting errors in the scientific literature and that results posted
to preprint servers were able to correct the record within weeks of the
initial publication while the traditional path took six months in both
cases. A final notable case where preprints have accelerated the
correction of the scientific record was a preprint posted by Bik et al.
reporting numerous cases of image manipulation in peer reviewed studies
(30). Their preprint was posted on April 20, 2016 and published in
\emph{mBio} on June 7, 2016 (31). Instead of using preprints to react to
published papers that have been through peer review, it would be
interesting to consider how the editorial process for these examples and
the infamous ``Arsenic Life'' paper (32) would have been different had
they initially been posted as preprints.

@@

As the total number of preprints has grown exponentially since the
creation of \emph{bioRxiv}, submission of microbiology-affiliated
preprints has largely followed this growth (\textbf{Figure 1A}).
Although preprints are still relatively new, the collection of
microbiology-affiliated preprints indicates widespread experimentation
with the format and considerable geographic diversity. Reflecting the
still relatively novelty of preprints, 1,132 (86.1\%) corresponding
authors who submitted a microbiology-affiliated preprint (N=1,314 total)
have posted a single preprint and 3.6\% have posted 3 or more preprints.
Corresponding authors that have posted microbiology-affiliated preprints
are from 60 countries and are primarily affiliated with institutions in
the United States (50.8\% of microbiology-affiliated preprints), United
Kingdom (11.9\%), and Germany (4.2\%). As the preprint format matures,
it will be interesting to see whether the fraction of authors that post
multiple preprints increases and whether the geographic diversity
amongst those authors is maintained.

As stated above, preprints offer researchers the opportunity to improve
the quality of their work by adding a more formal and public step to the
scientific process. Among the microbiology-affiliated preprints, 146
(9.3\%) had been commented on at least once and only 35 (2.2\%) more
than three times using the \emph{bioRxiv}-hosted commenting feature.
Although the hosted commenting is only one mechanism for peer review,
this result was somewhat disturbing since the preprint model implicitly
depends on people's willingness to offer others feedback. Regardless,
authors do appear to be incorporating feedback from colleagues or
editorial insights from journals as 404 (25.8\%) of preprints were
revised at least once. Among the preprints posted prior to January 1,
2016, 31.6\% of the Microbiology category preprints, 35.1\% of the
microbiology-affiliated preprints, and 33.8\% of all preprints have been
published. As noted above, not all authors submit their preprints to
journals. This would indicate that the ``acceptance rates'' are actually
higher. Regardless, considering that these acceptance rates are higher
than many peer-reviewed journals (e.g.~approximately 20\% at ASM
Journals), these results dispel the critique that preprints represent
overly preliminary research.

@@

This work was supported in part by funding to PDS from the National
Institutes of Health (P30DK034933). We appreciate the support of
Altmetric, Inc and Thompson Reuters who provided advanced programming
interface (API) access to their databases. The workflow utilized
commands in GNU make (v.3.81), GNU bash (v.4.1.2), and R (v.3.3.2).
Within R we utilized the cowplot (v.0.6.3), dplyr (v.0.5.0), ggplot2
(v.2.2.0.9000), httr (v.1.2.1), RCurl (v.1.95-4.8), rentrez (v.1.0.4),
rjson (v.0.2.15), rvest (v.0.3.2.9000), sportcolors (v.0.0.1), and tidyr
(v.0.6.0) packages. A reproducible version of this manuscript and
analysis is available at
\url{http://www.github.com/SchlossLab/Schloss_PrePrints_mBio_2017}.
